---
title: "Random Forests"
description: |
  A short description of the post.
author:
  - name: Wilson Sithole
    url: https://example.com
  - name: Cephas Svosve
    url: https://example.com
  - name: Blessing Nyatsanza
    url: https://example.com
date: 01-26-2021
output:
  distill::distill_article:
    self_contained: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(doParallel)
library(rpart)
library(caret)
library(rpart.plot)
library(randomForest)
library(GGally)
library(pROC)

```


Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo. Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt. Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem. Ut enim ad minima veniam, quis nostrum exercitationem ullam corporis suscipit laboriosam, nisi ut aliquid ex ea commodi consequatur? Quis autem vel eum iure reprehenderit qui in ea voluptate velit esse quam nihil molestiae consequatur, vel illum qui dolorem eum fugiat quo voluptas nulla pariatur?"

```{r}
#read in the data 
train <- read.csv("~/Desktop/Kaggle_Datasets/train.csv")
test <- read.csv("~/Desktop/Kaggle_Datasets/test.csv")
str(train)
str(test)
#is.na.data.frame(train)
```

# Exploratory Data Analysis and Cleaning

At vero eos et accusamus et iusto odio dignissimos ducimus qui blanditiis praesentium voluptatum deleniti atque corrupti quos dolores et quas molestias excepturi sint occaecati cupiditate non provident, similique sunt in culpa qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita distinctio. Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae. Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.

```{r}
#join train and test data
ltrain <- nrow(train)
new.test<- test%>%
          dplyr::mutate(Survived = "1") # to be deleted before final predictions are made 
full <- rbind(train, new.test) 

#change Pclass, Survived to factors 
full$Survived = factor(full$Survived)
full$Pclass = factor(full$Pclass) 

#missing values
sapply(full, function(x) sum(is.na(x)))
colSums(full == "")

#impute missing values 
full$Embarked[full$Embarked == ""] = "C"
mean_age <- mean(full$Age, na.rm = T)
sd_age <- sd(full$Age, na.rm = T) 
full$Age[is.na(full$Age)] = runif(1, mean_age - sd_age, mean_age + sd_age)
full$Fare[is.na(full$Fare)] = mean(full$Fare, na.rm = T)

```

Some plots to visualize the relationships between our label and the features
At vero eos et accusamus et iusto odio dignissimos ducimus qui blanditiis praesentium voluptatum deleniti atque corrupti quos dolores et quas molestias excepturi sint occaecati cupiditate non provident, similique sunt in culpa qui officia deserunt mollitia animi, id est laborum et dolorum fuga. Et harum quidem rerum facilis est et expedita distinctio. Nam libero tempore, cum soluta nobis est eligendi optio cumque nihil impedit quo minus id quod maxime placeat facere possimus, omnis voluptas assumenda est, omnis dolor repellendus. Temporibus autem quibusdam et aut officiis debitis aut rerum necessitatibus saepe eveniet ut et voluptates repudiandae sint et molestiae non recusandae. Itaque earum rerum hic tenetur a sapiente delectus, ut aut reiciendis voluptatibus maiores alias consequatur aut perferendis doloribus asperiores repellat.

## sex
```{r}
ggplot(full[1:ltrain,], aes(x = Sex, fill = Survived)) + geom_bar()
```

More females survived both number-wise and proportion-wise. 

## Pclass 
```{r}
ggplot(full[1:ltrain,], aes(x = Pclass, fill = Survived)) + geom_bar()
```

Class 1 had the highest proportion of survivors, followed by class 2 and lastly class 3. 

## Embarked 
```{r}
ggplot(full[1:ltrain,], aes(x = Embarked, fill = Survived)) + geom_bar()
```

The majority of passengers who embarked at station S did not survive. 

## Age
```{r}
ggplot(full[1:ltrain,], aes(x = Age, fill = Survived)) + geom_histogram()
```

Age seems like a good feature.Younger ages survived more. 

## SibSp
```{r}
ggplot(full[1:ltrain,], aes(x = SibSp, fill = Survived)) + geom_histogram()
```

Those with FEWER siblings + spouses tended to survive 

## Parch 
```{r}
ggplot(full[1:ltrain,], aes(x = Parch, fill = Survived)) + geom_histogram()
```


## Fare 

```{r}
ggplot(full[1:ltrain,], aes(x = Fare, fill = Survived)) + geom_histogram()
```

# Dealing with the names 

```{r}
officers <- c(" Capt",' Col',' Don',' Dr',' Jonkheer',' Major',' Rev',' Sir',' the Countess')
full <- full %>%
   mutate(Title_B = gsub('(.*,)|(\\..*)', '', Name))%>%
   mutate(Title = dplyr::case_when(
                      Title_B == " Mlle" ~ " Miss",
                      Title_B == " Ms" ~ " Miss",
                      Title_B == " Mme" ~ " Mrs",
                      Title_B == " Lady" ~ " Miss",
                      Title_B == " Dona" ~ " Miss",
                      Title_B %in% officers ~ "Officer",
                      TRUE ~ Title_B
  ))
full$Title = factor(full$Title)

#hist of titles 

ggplot(full[1:ltrain,], aes(x = Title , fill = Survived)) + geom_bar(position ="fill") + ylab("Frequency")

```

Ladies survived at a higher frequency than men. Masters survived at a higher rate than ordinary men.

# Modeling with rpart and caret 

```{r}
#creating train and test sets
train.new <- full[1:ltrain,]
ltest <- ltrain+1
test.new <- full[ltest: nrow(full),]
in.train <- createDataPartition(train.new$Survived, p=0.80,list=FALSE)
trainset <- train.new[in.train,]
testset <- train.new[-in.train,]

#parallel programming 
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

#fit the model 
set.seed(1)
model <- rpart(Survived ~ Pclass + Sex + Fare + Embarked + Title + Age,
               data = trainset,
               method = "class")
    
#stop clusters 
stopCluster(cl) 

#plot
rpart.plot(model) 
```

# Model Perfomance Metrics 

```{r}
predictions <- predict(model, testset, type = "class")
mean(predictions == testset$Survived) # accuracy score 

t2 <- table(predictions, testset$Survived) # contingency table 
t2 
precision <- t2[1,1]/sum(t2[1,])
recall <- t2[1,1]/sum(t2[,1])
F1_score <- 2*precision*recall/(precision + recall)
precision
recall
F1_score

#ROC

pROC_obj <- roc(testset$Survived, as.numeric(predictions),
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


sens.ci <- ci.se(pROC_obj)
plot(sens.ci, type="shape", col="lightblue")
plot(sens.ci, type="bars")

```


# Random forest 

```{r}
set.seed(3567)
#parallel 
cl <- makePSOCKcluster(8)
registerDoParallel(cl)

rf_model <- randomForest(Survived ~ Pclass + Sex + Fare + Embarked + Title + Age,
                         data = trainset,
                         importance = T,#performance boost
                         proximity = F,#performance boost
                         ntrees = 120, 
                         mtry = 3
                         )
#stop clusters 
stopCluster(cl) 

#Errors 
#plot(rf_model)
rf_model

```


```{r}
predictions_rf <- predict(rf_model, testset, type = "class")
mean(predictions_rf == testset$Survived) # accuracy score 

t2 <- table(predictions_rf, testset$Survived) # contingency table 
t2 
precision_rf <- t2[1,1]/sum(t2[1,])
recall_rf <- t2[1,1]/sum(t2[,1])
F1_score_rf <- 2*precision*recall/(precision + recall)
precision_rf
recall_rf
F1_score_rf

#ROC curve
library(pROC)
pROC_obj <- roc(testset$Survived, as.numeric(predictions_rf),
            smoothed = TRUE,
            # arguments for ci
            ci=TRUE, ci.alpha=0.9, stratified=FALSE,
            # arguments for plot
            plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
            print.auc=TRUE, show.thres=TRUE)


sens.ci <- ci.se(pROC_obj)
plot(sens.ci, type="shape", col="lightblue")
plot(sens.ci, type="bars")
```



